---
title: "Intro to R Lab"
author: "Ayodele Odubela"
date: "5/11/2018"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Introduction to R Lab: 

2.3.1- Basic Commands

In tis section the basic comamands of finding the length of vector as well as learning about (?) and adding data to a matrix. Here I also explore creating a matrix of normal random variables (mean of 0 and sd of 1) using the rnorm() function. I also set seeds for the random number functions in order to get reproducible random numbers. I also explored summary statistics for the variable y.  
```{r basic commands}
x <- c(1,2,3,5)
x
length(x)
y <- c(7,8,9)
x + y
ls()
rm(x,y)
rm(list=ls())
?matrix
x = matrix(data=c(1,2,3,4), nrow= 2, ncol = 2)
x
x = matrix(c(1,2,3,4), 2,2)
x
matrix(c(1,2,3,4), 2,2, byrow = TRUE)
sqrt(x)
x^2
x= rnorm(50)
y = x+ rnorm(50, mean = 50, sd= .1)
cor(x,y)


set.seed(1303)
rnorm(50)
set.seed(3)
y = rnorm(100)
mean(y)
var(y)
sqrt(y)
sd(y)
```


2.3.2 - Graphics

In this section I used the plot() function to display my matrix of random data. I used parameters xlab, ylab, and main to create labels for the plot as well as a title. I created a pdf and jpeg of my plot. I also used the seq() function to create a vector sequence. I implemented contour() to fine-tune my plot and persp() to make at 3D plot. 
```{r  graphics}
x = rnorm(100)
y = rnorm(100)
plot(x,y)
plot(x, y, xlab= "this is the x axis", ylab = "this is the y axis", main = "Plot of x and y")
pdf("Figure 1.pdf")
jpeg("Figure1j.jpeg")
plot(x,y,col = "green")
dev.off()

x =seq(1, 10)
x
x = 1:10
x

y = x
f = outer(x,y, function(x,y)cos(y)/(1+x^2))
contour(x,y,f)
contour(x,y,f, nlevels=45, add = T)
fa = (f-t(f))/2
contour(x,y,fa, nlevels = 15)


image(x, y, fa)
persp(x, y, fa)
persp(x, y, fa, theta = 30)
persp(x, y, fa, theta = 30, phi = 30)
persp(x, y, fa, theta = 30, phi = 70)
persp(x, y, fa, theta = 30, phi = 40)

```

2.3.3- Indexing Data

In this section I learned how to slice and index matrices with brackets, colons, and commas. This section on row/column selection I found important when it comes to selecting the RIGHT data. 
```{r}
A = matrix(1:16, 4, 4)
A

A[2,3]
A[c(1,3), c(2,4)]
A[1:3, 2:4]
A[1:2,]
A[,1:2]
A[1,]
A[-c(1,3) , ]
dim(A)
```


2.3.4 - Loading Data 

I am very familliar with loading data using functions like read.table() and read.csv() as well as their parameters for header and now to handle NA values. I used the mtcars data here which is preloaded into RStudio. 

2.3.5 - Graphical and Numerical summaries

Here I create a scatterplot of the cylinder and MPG data from the mtcars dataset. This plot shows clearly that the more cylinders a car has, the lower its MPG. I also created a boxplot using the same datapoints to better highlight the outliers that were recieving unusually high or low MPG. I created a bar chart and fine tuned its output graphics with additional parameters including x and y axis labels. The hist() function easily created a histogram and pairs() created a scatterplot matrix. I used the summary() function to show general summary of the dataset. The most important stats are included like min, max, median, and mean. 
```{r}
auto = mtcars
#fix(auto)
dim(auto)
auto = na.omit(auto)
names(auto)
plot(auto$cyl, auto$mpg)
attach(auto)
plot(cyl, mpg)
cyl= as.factor(cyl)
plot(cyl, mpg)
plot(cyl, col="red")
plot(cyl, col = "red", varwidth=T)
plot(cyl, col = "red", varwidth=T, horizontal = T)
plot(cyl, col = "red", varwidth=T, horizontal = T, xlab = "cylinders", ylab = "mpg")

hist(mpg, col = 2)
hist(mpg, col = 2, breaks = 15)

pairs(auto)
pairs(~mpg + disp + hp + wt, auto)
plot(hp, mpg)
identify(hp, mpg)

summary(auto)
summary(mpg)
```


