---
title: "<h3>MSDS660 - Week 1: Basic Concepts<br></h3>"
output:
  html_notebook: default
  pdf_document: default
---
---
<h4>Christy Pearson<br>May 12, 2018</h4>
<br>
<h5>This week's assignment is a Lab exercise to verify our R/RStudio setup and pratice working with R notebooks.</h5>
<br>
Preparing the current environment
```{r}
rm(list = ls()) # Prevents results from previous runs being carried over into new runs.
graphics.off()  # Clears the graphic plots window
cat("\014")     # Clears the console

setwd("C:/Users/Creat/OneDrive/Documents/MSDS660/Week 1")
getwd()
```
Package Installations
```{r}
# https://stackoverflow.com/questions/34739681/unable-to-move-temporary-installation-when-installing-dependency-packages-in-r/44256437#44256437
# Sometimes needed when "cannot move temporary installation" error
#debug(utils:::unpackPkgZip) 

# Packages for this rmd
# install.packages('sampling')
```
Lab Exercises:<br><br>
Assigning vectors
```{r}
# create a function of numbers
x <- c(1,2,3,5)
x
```
```{r}
# using '=' to store values rather than '<-'
x = c(1,6,2)
x
y = c(1,4,3)
y
```
Functions and Matrices
```{r}
# length() function
length(x)
length(y)
```
```{r}
# adding the two vectors togther
x+y
```
```{r}
# ls() and rm() functions
ls()
rm(x,y)
ls()
```
```{r}
# it is possible to combine ls() and rm()
x = c(1,6,2)
x
y = c(1,4,3)
y
rm(list=ls())
ls()
```
```{r}
# matrix() function
?matrix  #opens up the R documentation within RStudio

# time to create a matrix
x = matrix(data=c(1,2,3,4), nrow=2,ncol=2)
x
```
```{r}
# another way to create the same matrix
x = matrix(c(1,2,3,4),2,2)
x
```
```{r}
# add the byrow operator to load the matrix based on rows and not columns
y <- matrix(c(1,2,3,4), 2, 2, byrow=TRUE)
y
```
Notice the difference between matrix x and y.(Hint: look at where values '2' and'3' are placed)<br><br>
Statistical Functions
```{r}
# sqrt() and exponent functions
sqrt(x)
x^2

```
Note: we didn't reassign the values of x, we just used them in our calculations
```{r}
# rnorm() and cor() functions
# rnorm()- generates a vector of random normal variables
x = rnorm(50)
# cor() - correlated set of numbers
y=x+rnorm(50, mean=50, sd=.1)
cor(x,y)
```
```{r}
# use set.seed whenever performing calculations involving random quantities
set.seed(1303)
rnorm(50)
```
```{r}
# functions mean(), var(), sd()
set.seed(3)
y=rnorm(100)
mean(y)
```
```{r}
var(y)
```
```{r}
sqrt(var(y))
```
```{r}
sd(y)
```
Graphics
```{r}
# plot()
x = rnorm(100)
y = rnorm(100)
plot(x,y)
```
```{r}
# Now let's make is pretty
plot(x,y,xlab="x-axis",ylab="y-axis",main="Plot of X vs Y")
```
```{r}
# saving the output of a R plot: functions pdf() and jpeg()
pdf("Figure.pdf")
plot(x,y,col="green")
dev.off()
```
The pdf was written tot he wroking directory that we specified above. Also note the dev.off() is a way to tell R that we are done with the plot()
```{r}
# seq() function
x = seq(1,10)
x
```
```{r}
# here's another way
x=1:10
x
```
```{r}
# fancier sequence
x = seq(-pi,pi,length=50)
x
```
```{r}
# time to create some sophisticated plots with all of this
# contour() function - topographical map plotting with three arguments
# 1) a vector of x values
# 2) a vector of y values
# 3) a matrix whose elements correspond to the z value for each (x,y)
y=x
f=outer(x,y,function(x,y)cos(y)/(1+x^2))
contour(x,y,f)
contour(x,y,f,nlevels=45,add=T)
```
```{r}
# another contour plot
fa=(f-t(f))/2
contour(x,y,fa,nlevels=15)
```
```{r}
# function image()- works the same as contour() except it produces a heat map.
# a heat map is a plot where the z value is represented by a code coded plot
image(x,y,fa)
```
```{r}
# function persp() - used to produce a three dimensional map
persp(x,y,fa)
```
```{r}
# slightly different view
persp(x,y,fa,theta=30)
```
```{r}
# slightly different view
persp(x,y,fa,theta=30,phi=20)
```
```{r}
# slightly different view
persp(x,y,fa,theta=30,phi=70)
```
```{r}
# slightly different view
persp(x,y,fa,theta=30,phi=40)
```
Indexing Data
```{r}
# indexing data
A = matrix(1:16,4,4)
A
```
```{r}
# retrieve one element of A
A[2,3]
```
```{r}
# select a sub matrix
A[c(1,3),c(2,4)]
```
```{r}
# another sub matrix
A[1:3,2:4]
```
```{r}
# another sub matrix - specific rows
A[1:2,]
```
```{r}
# another sub matrix - specific columns
A[,1:2]
```
```{r}
# another sub matrix
A[1,]
```
```{r}
# another sub matrix using the '-' command
A[-c(1,3),]
```
```{r}
# last of the matrices
A[-c(1,3),-c(1,3,4)]
```
```{r}
# dim() function
dim(A)
```
Loading Data
```{r}
# load the Auto.data file
Auto=read.table("http://www-bcf.usc.edu/~gareth/ISL/Auto.data")
fix(Auto)
```
The output from the fix() command opens in a seperate window. Not sure how to make it open in here.
```{r}
# what about the header
Auto=read.table("http://www-bcf.usc.edu/~gareth/ISL/Auto.data", header=T, na.strings="?")
fix(Auto)
```
The output from the fix() command opens in a seperate window. Not sure how to make it open in here.
```{r}
# loading from the csv
Auto=read.csv("http://www-bcf.usc.edu/~gareth/ISL/Auto.csv", header=T, na.strings="?")
fix(Auto)
```
The output from the fix() command opens in a seperate window. Not sure how to make it open in here.

```{r}
# look at some of the data particulars
dim(Auto)
Auto[1:4,]
```
```{r}
# na.omit() - handle missing observations
Auto=na.omit(Auto)
dim(Auto)
```
cleaned up 5 rows of missing data
```{r}
# column names
names(Auto)
```
Graphical and nemerical Summaries
```{r}
# plot the cylinder vs mpg of the Auto dataset
plot(cylinders, mpg)
```
The error is because R doesn't know where to find cylinders and mpg. Need to specify the dataset
```{r}
# plot the cylinder vs mpg of the Auto dataset
plot(Auto$cylinders, Auto$mpg)
```
```{r}
# alternatively, use the attach() function
attach(Auto)
plot(cylinders, mpg)
```
```{r}
# convert cylinder from vector to factor
cylinders <- as.factor(cylinders)

# now plot cylinders vs mpg
plot(cylinders, mpg)
```
```{r}
# another plot
plot(cylinders, mpg, col="red")
```
```{r}
# another plot
plot(cylinders, mpg, col="red",varwidth=T)
```
```{r}
# another plot
plot(cylinders, mpg, col="red",varwidth=T,horizontal=T)
```
```{r}
# another plot
plot(cylinders, mpg, col="red",varwidth=T, xlab="cylinders", ylab="mpg")
```
```{r}
# hist() to make a histogram
hist(mpg)
```
```{r}
# another histogram
hist(mpg,col=2)
```
```{r}
# another histogram
hist(mpg,col=2,breaks=15)
```
```{r}
# paris() function
pairs(Auto)
```
Those are hard to read
```{r}
# pairs() one more time
pairs(~ mpg+displacement+horsepower+weight+acceleration,Auto)
```
That's a little better
```{r}
# identify() function - will identify the value of a given point
plot(horsepower,mpg)
identify(horsepower,mpg,name)
```
I couldn't get the identify() to work in the notebook.   However, I did in an R script.   I'll attach that the the project submission too.
```{r}
# summary() function
summary(Auto)
```
```{r}
# for one element
summary(mpg)
```
```{r}
# other commands
#savehistory()
#loadhistory()
```

