---
title: "Meng Luo<br>
MSDS - 660 Statistical Methods and<br> 
Experimental Design<br>"
output: html_notebook
---
##Week 2: Simple Linear Regression - Exercise: Linear Regression Using R
May 23rd, 2018
```{r,echo=FALSE,results=FALSE}
rm(list=ls()) # Removes all objects to prevent results from previous runs being carried over into new runs.
graphics.off()  # Clears the graphic plots window
cat("\014")     # Clears the console
# setwd("C:/Codes/Data/MSDS660Stats/Week2")
library(MASS)# https://cran.r-project.org/web/packages/MASS/MASS.pdf
```
###1. Data set description
We will use the "Housing Values in Suburbs of Boston" data set from MASS package version 7.3-50. This data frame was collected back in 1978 and it seems pretty clean with 506 observations and 14 columns. The columns are:

crim: per capita crime rate by town.<br>
zn: proportion of residential land zoned for lots over 25,000 sq.ft.<br>
indus:  proportion of non-retail business acres per town.<br>
chas: Charles River dummy variable (= 1 if tract bounds river; 0 otherwise).<br>
nox:  nitrogen oxides concentration (parts per 10 million).<br>
rm: average number of rooms per dwelling.<br>
age:  proportion of owner-occupied units built prior to 1940.<br>
dis:  weighted mean of distances to five Boston employment centres.<br>
rad:  index of accessibility to radial highways.<br>
tax:  full-value property-tax rate per \$10,000.<br>
ptratio:  pupil-teacher ratio by town.<br>
black:  sqr(1000(Bk-0.63)) where Bk is the proportion of blacks by town.<br>
lstat:  lower status of the population (percent).<br>
medv: median value of owner-occupied homes in \$1000s.<br>

(https://cran.r-project.org/web/packages/MASS/MASS.pdf)

```{r}
head(Boston)
tail(Boston)
```
###2. Exploratory data analysis
```{r}
Boston_4 <- subset(Boston, select = c( "rm","medv", "dis", "indus"))
plot(Boston_4)
```
We can find the data pairs of (indus,dis) and (rm, medv) showing certain patterns. However, before we go further into calculation, there are some out-lined records with rm values of 50. Let's create a data subset 'BostonX' without them and see how it will affect our results. 
```{r}
Boston[order(Boston$medv,decreasing=T)[1:20],]
summary(Boston$medv)
BostonX <- subset(Boston, medv<50)
summary(BostonX$medv)
plot(density(Boston$medv))
plot(density(BostonX$medv))
```

I am not sure whether these points with medv= 50 are outliners: they may be placeholders for missing data or legit data simply because there were many properties prized at $50,000 back then in Boston. We will calculate without outliners first and the other scenario later to see what difference it makes.

###3. Data pairs correlation
According to the scatterplots we draw in the last section for the data subset with 4 columns of rm, medv, dis, and indus, the pair (rm, medv) shows a positive correlation while (dis, indus) shows a negative one. And (rm, indus) has no correlation.

```{r}
BostonX_2a <- subset(BostonX, select = c( "rm", "medv"))
plot(BostonX_2a)
BostonX_2b <- subset(BostonX, select = c( "indus", "dis"))
plot(BostonX_2b)
BostonX_2c <- subset(BostonX, select = c( "rm", "indus"))
plot(BostonX_2c)
```

###4,5,6. Linear regression on data pair (rm, medv)
By choosing the variables pairs of average room number(rm) and median value of homes (medv), we try to decide whether they have correlation.<br>
The null hypotheis H0="rm and medv have no correlation".(beta = 0)
 <br>
The alternative hypotheis H1="rm and medv have  correlation".(beta <> 0) <br>
```{r}
plot(BostonX_2a, sub="without outliners")
BostonX.lm <- lm(rm ~ medv, BostonX_2a)
summary(BostonX.lm)
```
###7,8. Result discussion
The result shows rm and medv have a p-value lower than 5% significance level for the null hypotheis. Therefore there is evidence of a significant relationship and our H0 is rejected. It proves that rm and medv have correlation with linear relationship. <br>

###9. Model assumption verification
Linearity: we can verify it by observing the rm vs medv plot. The low p-value also shows evidence in our linear model. <br>
Normality: The residuals have normal distribution around zero and the mean of residuals is almost zero. Therefore the assumption holds true.

```{r}
plot(density(BostonX.lm$residuals))
mean(BostonX.lm$residuals)
```

Constant variance of residuals: We can confirm it by plotting the residuals, which is randomly distributed around zero.
```{r}
plot(BostonX.lm$residuals)
```

In addition, by running a command for correlation coefficient, we can confirm that the model is a good fit.

```{r}
 cor(BostonX_2a,use = "all.obs",method = "pearson") 
```
### 10 Conclusion: Findings and concerns
The calulation shows that values of rm and medv have a position correlation. If we investigate further, the existance of outliners in the original dataset will not hurt our model and assuptions.
```{r}
Boston_2a <- subset(Boston, select = c( "rm", "medv"))
plot(Boston_2a, sub="with outliners")
Boston.lm <- lm(rm ~ medv, Boston_2a)
summary(Boston.lm)
cor(Boston_2a,use = "all.obs",method = "pearson") 
```